Appendix 4: species distribution model comparison between inla and sdmTMB

### 2021-12-16 doc/appendix-binomial-inla.Rmd
add a couple more years to analysis

```{r}
d <- pcod %>% filter(year >= 2011)
yr_lu <- data.frame(year = unique(d$year), i_year = seq_along(unique(d$year)))
d <- dplyr::left_join(d, yr_lu)
```

@@

```{r , out.width="4in"}
plot(sdmtmb_coefs$estimate, inla_coefs$mean)
abline(a = 0, b = 1)
```

@@

post_sdmTMB2 <- post_sdmTMB %>% mutate(variable = factor(`.variable`, 
                    levels = unique(post_sdmTMB$`.variable`), 
                    labels = c( "2011", "2013", "2015", "2017", 
                                "Range", "AR1 rho", "Random Field SD")))
```

@@

```{r inla-densities}
# post_inla <- inla.posterior.sample(result = m_inla, n = 400L)
x <- list()
x[[1]] <- inla.smarginal(m_inla$marginals.fixed$year_factor2011) %>% 
  as_tibble() %>% mutate(variable = "as.factor.year.2011") 
x[[2]] <- inla.smarginal(m_inla$marginals.fixed$year_factor2013) %>% 
  as_tibble() %>% mutate(variable = "as.factor.year.2013") 
x[[3]] <- inla.smarginal(m_inla$marginals.fixed$year_factor2015) %>% 
  as_tibble() %>% mutate(variable = "as.factor.year.2015") 
x[[4]] <- inla.smarginal(m_inla$marginals.fixed$year_factor2017) %>% 
  as_tibble() %>% mutate(variable = "as.factor.year.2017") 
x[[5]] <- inla.smarginal(m_inla$marginals.hyperpar$`Range for i`) %>% 
  as_tibble() %>% mutate(variable = "range") 
x[[6]] <- inla.smarginal(m_inla$marginals.hyperpar$`Stdev for i`) %>% 
  as_tibble() %>% mutate(variable = "sigma_E") 
x[[7]] <- inla.smarginal(m_inla$marginals.hyperpar$`GroupRho for i`) %>% 
  as_tibble() %>% mutate(variable = "ar1_rho") 
post_inla1 <- do.call(rbind, x)

post_inla2 <- post_inla1 %>% mutate(variable = factor(variable, 
                    levels = unique(post_inla1$variable), 
                    labels = c( "2011", "2013", "2015", "2017", 
                                "Range", "Random Field SD", "AR1 rho")))
```

@@

```{r sdm-predictions}
nd <- filter(qcs_grid, year >= 2011)
pred <- predict(m, newdata = nd) 
```

```{r maps, cache=FALSE, out.width="4in"}
ggplot(pred) + geom_raster(aes(X, Y, fill = plogis(est))) + 
  facet_wrap(~year) +   scale_fill_viridis_c() + coord_fixed() 
```

@@

```{r inla-pred, out.width="4in"}
idx <- inla.stack.index(sdat, 'stdata')$data
d$inla_pred <- m_inla$summary.fitted.values[idx, "mean"]
ggplot(d) + geom_point(aes(X, Y, colour = inla_pred), size = 2) + 
  facet_wrap(~year) + scale_colour_viridis_c() + coord_fixed() 
```

>>>>

```{r}
d <- pcod %>% filter(year >= 2007)
yr_lu <- data.frame(year = unique(d$year), i_year = seq_along(unique(d$year)))
d <- dplyr::left_join(d, yr_lu)
```

@@

```{r , out.width="5.5in"}
plot(sdmtmb_coefs$estimate, inla_coefs$mean)
abline(a = 0, b = 1)
```

@@

post_sdmTMB2 <- post_sdmTMB %>% mutate(variable = factor(`.variable`, 
                  levels = unique(post_sdmTMB$`.variable`), 
                  labels = c( "2007", "2009", "2011", "2013", "2015", "2017", 
                                "Range", "AR1 rho", "Random Field SD")))
```

@@

```{r inla-densities}
# post_inla <- inla.posterior.sample(result = m_inla, n = 400L)
x <- list()
x[[1]] <- inla.smarginal(m_inla$marginals.fixed$year_factor2007) %>% 
  as_tibble() %>% mutate(variable = "as.factor.year.2007") 
x[[2]] <- inla.smarginal(m_inla$marginals.fixed$year_factor2009) %>% 
  as_tibble() %>% mutate(variable = "as.factor.year.2009") 
x[[3]] <- inla.smarginal(m_inla$marginals.fixed$year_factor2011) %>% 
  as_tibble() %>% mutate(variable = "as.factor.year.2011") 
x[[4]] <- inla.smarginal(m_inla$marginals.fixed$year_factor2013) %>% 
  as_tibble() %>% mutate(variable = "as.factor.year.2013") 
x[[5]] <- inla.smarginal(m_inla$marginals.fixed$year_factor2015) %>% 
  as_tibble() %>% mutate(variable = "as.factor.year.2015") 
x[[6]] <- inla.smarginal(m_inla$marginals.fixed$year_factor2017) %>% 
  as_tibble() %>% mutate(variable = "as.factor.year.2017") 
x[[7]] <- inla.smarginal(m_inla$marginals.hyperpar$`Range for i`) %>% 
  as_tibble() %>% mutate(variable = "range") 
x[[8]] <- inla.smarginal(m_inla$marginals.hyperpar$`Stdev for i`) %>% 
  as_tibble() %>% mutate(variable = "sigma_E") 
x[[9]] <- inla.smarginal(m_inla$marginals.hyperpar$`GroupRho for i`) %>% 
  as_tibble() %>% mutate(variable = "ar1_rho") 
post_inla1 <- do.call(rbind, x)

post_inla2 <- post_inla1 %>% mutate(variable = factor(variable, 
                  levels = unique(post_inla1$variable), 
                  labels = c( "2007", "2009", "2011", "2013", "2015", "2017", 
                                "Range", "Random Field SD", "AR1 rho")))
```

@@

```{r sdm-predictions}
nd <- filter(qcs_grid, year >= 2007)
pred <- predict(m, newdata = nd) 
```

```{r maps, cache=FALSE, out.width="6in"}
ggplot(pred) + geom_raster(aes(X, Y, fill = plogis(est))) + 
  facet_wrap(~year) + scale_fill_viridis_c() + coord_fixed() 
```

@@

```{r inla-pred, out.width="6in"}
idx <- inla.stack.index(sdat, 'stdata')$data
d$inla_pred <- m_inla$summary.fitted.values[idx, "mean"]
ggplot(d) + geom_point(aes(X, Y, colour = inla_pred), size = 2) + 
  facet_wrap(~year) + scale_colour_viridis_c() + coord_fixed() 
```
